---
title: "Colleges"
author: "Oluchi Nnadi"
format: html
execute:
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(primer.data)
library(tidymodels)
library(broom)
library(marginaleffects)
```



```{r}
colleges |> 
  ggplot(aes(x = tuition, y = grad_rate)) +
  geom_point(alpha = 0.4) +
  geom_smooth(method = "loess", formula = y ~ x, se = FALSE) +
  labs(
    title = "Graduation Rate vs. Tuition",
    subtitle = "Higher tuition is associated with higher graduation rates",
    x = "Tuition (in $10,000s)",
    y = "Graduation Rate"
  )
```

```{r}
#| cache: true
x <- colleges |> 
  select(tuition, grad_rate, selectivity) |> 
  filter(tuition > 2) |> 
  drop_na() |> 
  mutate(selectivity = factor(
    selectivity, 
    levels = c("Elite", "Highly Selective", "Moderately Selective", "Lowly Selective", "Non-selective")
  ))
```

```{r}
#| cache: true
fit_colleges <- linear_reg() |> 
set_engine("lm") |>
 fit(grad_rate ~ tuition + selectivity, data = x)
```

```{r}
tidy(fit_colleges, conf.int = TRUE)
```

Summary: For a college/university, graduation rates can be a very important indicator of it's quality of education and demographic of students, but how does tutition impact this measure? Using data from IPEDS on U.S. 4-year colleges and universities, we plan to build a model that will measure the difference in expected graduation rates between colleges with tuition of $20,000 and colleges with tuition of $30,000. However, our model could have some limitations due to tuition not necessarily being representative of the amount students actually have to pay because financial aid varies from school to school; this could cause it's influence on graduation rates to change. We modeled the normally distributed variable of graduation rate as a linear function of tuition and selectivity.
